### 2025-01-11 msred-short.tex

%+----------------------------------------------------------------------------+
%| SLIDES: Multisymplectic geometry, moment maps and reduction.
%| Contents:	- 25 minutes (extimated duration ~2 minutes per slide, ~12 slides )
%|
%| Author: Antonio miti
%| Event: 17th International Young Researchers Workshop on Geometry, Mechanics and Control
%| Place: Leuven
%| Date: 28/3/23
%+----------------------------------------------------------------------------+

%- HandOut Flag -----------------------------------------------------------------------------------------
	\newif\ifHandout
	%\Handouttrue  %uncomment for the printable version
	%Handling of flags it is not preserved when passing to standalone-subfiles!

%- D0cum3nt ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{./custom-style}
\usepackage{./math}
\usetikzlibrary{positioning}
\usepackage{multicol}
\usepackage{xfrac}

\renewcommand{\action}{\curvearrowright} 
\makeatletter
\def\blfootnote{\gdef\@thefnmark{}\@footnotetext}
\makeatother

\newcommand{\fgmodule}{\mathfrak{X}_{\g}}
\newcommand{\tgtvanform}{I_{\Omega}(N)}
\newcommand{\tgtvf}{\mathfrak X_N(M)}
\newcommand{\vanvf}{I_{\mathfrak{X}}(N)}%{I_N^{\mathfrak X}}
\newcommand{\Ham}{\mathsf{Ham}}
\renewcommand{\ham}{\text{ham}}

%--Beamer Style-----------------------------------------------------------------------------------------------
\usetheme{toninus}

\usetikzlibrary{backgrounds}
  \tikzset{
    invisible/.style={opacity=0},
    visible on/.style={alt=#1{}{invisible}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
  }

%- T1tle P4g3 -------------------------------------------------------------------------------------------
\title{ Symmetries and Reduction \\ of Multisymplectic Manifolds } 
%\subtitle{\href{https://www.mpim-bonn.mpg.de/node/158}{MPI-Oberseminar}}
\subtitle{\href{https://wis.kuleuven.be/events/young-researchers-workshop2023/young-researchers-workshop2023}{17th International Young Researchers Workshop on Geometry, Mechanics and Control}}
\author[AMM]{\href{https://dmf.unicatt.it/miti/}{Antonio Michele Miti}}

\institute[Mpim]{
	Università Cattolica del Sacro Cuore, Brescia, Italy 
	\\
	\vspace{.5em}
	\href{https://dipartimenti.unicatt.it/dmf-home?rdeLocaleAttr=it}{\includegraphics[width=3.5cm]{Logos/UnicattBS-logo}} 
}
\date[YGMC] % (optional, should be abbreviation of conference name)
{	
	{\vskip 1ex}
	%Bonn, April 7, 2022
	KU Leuven, Belgium, March 2023
}

%---------------------------------------------------------------------------------------------------------------------------------------------------
%- D0cum3nt ----------------------------------------------------------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
	\end{frame}
	\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
%
Multisymplectic manifolds are a straightforward generalization of symplectic manifolds where closed non-degenerate k-forms are considered in place of 2-forms.
A natural theme that arises when dealing with (multi)symplectic structures is investigating the relationship between symmetries (group action preserving the fixed differential form) and reduction.

In this context, a reduction scheme can be taught as a procedure to construct a second space of reduced dimension out of a given (multi)-symplectic manifold with symmetries. The so obtained reduced space enjoys the convenient property of embodying the relevant geometric structure of the starting unreduced object.

A well-known result in symplectic geometry, known as Marsden–Weinstein–Meyer theorem, states that the relevant geometric structure of a symplectic manifold can be studied on the orbits contained in a regular level set of a so-called "momentum map".
Sniaticki and Weinstein have further extended this result to encompass singular momentum maps.

The scope of this talk is to review some relevant algebraic structures related to multisymplectic manifolds, namely the higher version of the observables algebra and moment maps, and discuss how the regular and singular reduction schemes extend to the multisymplectic framework.

This talk is based on ongoing joint work with Casey Blacker and Leonid Ryvkin.

}
%---------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Keywords}
\tikzstyle{every picture}+=[remember picture]
	\begin{columns}
    	\begin{column}{.45\textwidth}
    		\onslide<4->{
			\tikz[baseline]{
		            \node[draw=orange!40,anchor=base,text width=5cm] (s1)
		            {Group of transformations preserving the prescribed geometric structures.};
			}
		}
	\end{column}
    	\begin{column}{.45\textwidth}
    		\onslide<5->{
			 \tikz[baseline]{
		            \node[draw=blue!40,anchor=base, text width=5cm] (s2)
		            {Procedure providing another manifold of reduced dimension encoding the relevant geometrical information.};
			}
		}
	\end{column}
	\end{columns}

	\vfill

	\begin{center}
		\large
		 \tikz[baseline]{
		            \node[fill=orange!20,anchor=base] (t1)
		            {Symmetries};
			}
		 and
		 \tikz[baseline]{
		            \node[fill=blue!20,anchor=base] (t2)
		            {Reduction};
		        } 
		\\
		of 
		 \tikz[baseline]{
	            \node[fill=green!20,anchor=base] (t3)
	            {Multi};
		}
		\kern-3pt - \kern-3pt
		 \tikz[baseline]{
	            \node[fill=red!20,anchor=base] (t4)
	            {Symplectic};
		}		
        manifolds
	\end{center}

	\vfill

	\begin{columns}
    	\begin{column}{.45\textwidth}
    		\onslide<3->{
	 		 \tikz[baseline]{
	            \node[draw=green!40,anchor=base,text width=5cm] (s3)
	            {A certain higher version \\(involving differential forms in degree $\geq 2$)};
	         }
		}		   	
		\end{column}
    	\begin{column}{.45\textwidth}
    		\onslide<2->{
				\tikz[baseline]{
	            \node[draw=red!40,anchor=base,text width=5cm] (s4)
	            {Geometric structure providing a prescription on how to measure the area of 2-dimensional surface elements embedded in the manifold.};
	           }	
			}
		\end{column}
	\end{columns}

	\begin{tikzpicture}[overlay]
        \path[->,draw=orange!40]<4-> (s1) edge [bend right] (t1);
        \path[->,draw=blue!40]<5-> (s2) edge [bend left] (t2);
        \path[->,draw=green!40]<3-> (s3) edge [bend left] (t3);
        \path[->,draw=red!40]<2-> (s4) edge [bend right] (t4);
	\end{tikzpicture}

	\vfill
	\onslide<1->{
	\begin{block}{Based on:}
		 \emph{ \small
			 Blacker, M, Ryvkin;
			\textbf{Reduction of $L_\infty$-algebra of observables on multisymplectic manifolds}; 
			\href{https://arxiv.org/abs/2206.03137}{arXiv:2206.03137}
		}
	\end{block}
	}

\end{frame}
\note[itemize]{
	\item Provide a vague idea of the words that make up the title.
	\item
	\item Conventions:
	\\- $M$ and $G$ are connected,
	\\- actions $\theta:G \curvearrowright M$ are always smooth
	\\- $\xi,\eta\in\g$,
	\\- for $\mu\in\Omega^*(M,\g^*)$ and $\xi\in\g$, write
			\[
				\mu_\xi := \langle\mu,\xi\rangle \;{\color{black!50}\in\Omega^*(M)}
			\]
			for the ``$\xi$th component'' of $\mu$.
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Symplectic Reduction}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Symplectic geometry (mechanics flavour)}
\begin{columns}[T]
	\begin{column}{.50\linewidth}
		\centering
		\textit{ "geometric approach" to mechanics \dots}
		%
		\begin{columns}
			\begin{column}{.60\linewidth}
				\begin{center}
					\includegraphics[width=0.6\linewidth]{Pictures/pendulum13}			
				\end{center}
			\end{column}	
			\begin{column}{.40\linewidth}
				\begin{center}
					\includegraphics[width=0.45\linewidth]{Pictures/pendulum-phase-space}			
				\end{center}
			\end{column}	
		\end{columns}
		%
		\begin{defblock}[Symplectic Manifold]
			\vspace{-1em}
			\includestandalone[width=1\textwidth]{Pictures/Figure_sym}	
		\end{defblock}
		%
		\pause
		\begin{exblock}[$M = T^\ast Q$ is symplectic]
			with $\omega = d \theta $ given by
			$$ \left.\theta\right\vert_{(q,p)} (v) = p (\pi_\ast v) ~.$$
		\end{exblock}
		%
		\pause
		\vspace{1.2em}
		\centering
		\textit{ based on the notion of \\"states".}		
	\end{column}
	\onslide<1->{\vrule{}}
	\pause
	\begin{column}{.50\linewidth}
		\centering
		\textit{ "algebraic approach" to mechanics \dots}
		\vspace{.5em}	
		\begin{defblock}[Classical Observables]
			Unital, associative, commutative algebra $C^\infty(M)$.
		\end{defblock}
		%
		\vspace{.1em}
		\pause
		\begin{defblock}[Hamiltonian vector fields]
			$\vHam_f \in \mathfrak{X}(M)$ such that:
			$$\iota_{\vHam_f} \omega = -df \quad$$ %$\in B^1(M)$
			%
			\footnotetext{	$\vHam_f$ = \emph{Ham.v.f. pertaining to $f\in C^\infty(M)$}.}
		\end{defblock}
		\vspace{.1em}
		%
		\begin{defblock}[Poisson Algebra of Observables]
			$C^\infty(M)$ is a Poisson algebra with
			$$\{f,g\} = \iota_{\vHam_g} \iota_{\vHam_f} \omega = \omega(\vHam_f,\vHam_g) ~.$$
		\end{defblock}
		%
		\pause
		\vspace{.15em}
		\centering
		\textit{ based on the notion of \\"measurable quantities".}						
	\end{column}
\end{columns}
\end{frame}
\note[itemize]{
	\footnotesize

	\item We work in the framework of multisymplectic geometry which is one of the possible generalizations of the well-established field of symplectic geometry.

	\item To recall what symplectic geometry is let me assume a particular point of view: mechanics.
	\\
	Idea:"
	Symplectic geometry is a branch of differential geometry studying symplectic manifolds; it originated as a formalization of the mathematical apparatus of classical mechanics and geometric optics."{\href{https://ncatlab.org/nlab/show/symplectic+geometry}{nlab}}

	Namely, a sym. mfd. is the geometric structure encoding the phase space of conservative, ordinary, classical, mechanical systems.

	\item $\theta$ = \emph{tautological 1-form}.
		$\theta$ evaluated at $p\in T^*Q$ in the fibre of $q\in Q$ and contracted with $v$ coincides with the form $p$ evaluated at $q$ and contracted with the push forward of $v$.

	\item We identify a special class of vector fields.
		Out of them one can define a Lie bracket.

	\item Poisson is a Lie algebra with the extra property of compatibility with the associative product (Leibniz rule)

	\item take away message: geometric (based on "states") vs algebraic (based on "measurable quantities").7
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\begin{frame}{Reminder: momentum maps in symplectic geometry}\label{frame:symplecticmomaps}
	Consider $\theta:G\curvearrowright M$ symplectic,~~ $\underline{\cdot}:\mathfrak{g}\to \mathfrak{X}(M)$ infinitesimal action.
	\vfill

	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			\onslide<2->{
				\begin{defblock}[Equivariant moment map]
					Smooth map $$\mu:M\to \mathfrak{g}^\ast$$
					such that:
					\begin{itemize}
						\item[i.] $d\langle \mu,\xi\rangle = -\iota_{\underline{\xi}}\omega$ 
						~\qquad, $\forall \xi \in \mathfrak{g}$
						\item[ii.] $\mu \circ \theta_g = Ad_g^\ast \circ \mu$
						 \qquad\small, $\forall g \in G$
					\end{itemize}
				\end{defblock}
			}
		\end{column}	
		%
		\onslide<2->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\onslide<3->{			
				\begin{defblock}[Comoment map]
					Linear map $$\widetilde{\mu}: \mathfrak{g}\to C^\infty(M,\omega)$$
					such that:
					\begin{itemize}
						\item[i.] $d\widetilde{\mu}(\xi) = -\iota_{\underline{\xi}}\omega$ 
						\qquad~\small, $\forall \xi \in \mathfrak{g}$
						\item[ii.] $\widetilde{\mu}([\xi,\eta]) = \lbrace\widetilde{\mu}(\xi),\widetilde{\mu}(\eta)\rbrace$ \small, $\forall \xi,\eta \in \mathfrak{g}$
					\end{itemize}
				\end{defblock}
			}
		\end{column}	
	\end{columns}	
	%
	\pause
	\vspace{1em}
	%
	\begin{columns}[]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.40\linewidth}
			%
			\centering \it
			\onslide<5->{
				\begin{upshotblocktitle}[Duality]
					\begin{displaymath}
						\mu(x) : \mathfrak{\xi} \mapsto \widetilde{\mu}(\xi)\big\vert_x
					\end{displaymath}
					%
					\emph{
					\small
					"duality wrt. the currying operation"					
					}
				\end{upshotblocktitle}
			}
		\end{column}	
		%
		%
		\begin{column}{.60\linewidth}
			\onslide<4->{			
				\begin{upshotblocktitle}[$\widetilde{\mu}$ as a lift]
					\begin{displaymath}
						\begin{tikzcd}[ampersand replacement = \&]
						 \& C^\infty(M,\omega) \ar[d,"\vHam"]
						 \\
						 \mathfrak{g} \ar[ur,dashed,sloped,"\widetilde{\mu}"]\ar[r,"\underline{\cdot}"'] \& \mathfrak{X}(M)
						\end{tikzcd}
					\end{displaymath}
					%
					\emph{
					\small
					"it is a lift (in the Lie category) of the infinitesimal action by the assigment of hamiltonian v.fields."					
					}
				\end{upshotblocktitle}
			}
		\end{column}	
	\end{columns}		
\end{frame}
\note[itemize]{
	\item consider an action preserving the symplectic form.
	\item[-] $\langle \cdot, \cdot \rangle$ is the natural pairing of $\mathfrak{g}$ and $\mathfrak{g}^\ast$.
		\item[-] $\theta_g$ is the diffeomorphism associated to $g\in G$ via $\theta:G\curvearrowright M$
		\item[-] $Ad^\ast_g$ is the coadjoint action $G\curvearrowright \mathfrak{g}^\ast$
	\item The moment map can be reexpressed as a comomentum map.
	\item Observe that ii. on the left always implies ii. on the right. The converse is trure only if $G$ is connected.
	\item $G\curvearrowright M$ is said "Hamiltonian" (see slide \ref{frame:gisthamaction} in appendix) iff $\exists$ comoment map $\widetilde{\mu}$.
	\item \emph{moment map: Describe ${\color{orange}G\curvearrowright M}$ in terms of $\color{blue}C^\infty(M)\to\X(M)$.}
	\begin{center}
	\begin{tikzpicture}[scale=1.5]
		\node[orange] (A) at (0,0) {$\g$};
		\node[gray] (A1) at (0,-.3) {$\xi$};
		\node[blue] (B) at (1,1) {$C^\infty(M)$};
		\node[gray] (B2) at (1.55,1) {$f$};
		\node[orange] (C) at (1,0) {$\X(M)$};
		\node[gray] (C1) at (1,-.3) {$\underline\xi$};
		\node[gray] (C2) at (1.55,0) {$X_f$};

		\path[->,blue,dashed] (A) edge node[above left] {$\langle\mu,\cdot\,\rangle$} (B);
		\path[->,orange] (A) edge (C);
		\path[->] (B) edge (C);
		\path[|->,gray] (A1) edge (C1);
		\path[|->,gray] (B2) edge (C2);

		\begin{scope}[xshift=3cm]
			\node at (-.32,.8) {as Lie algebras};
			\node at (-.13,.2) {i.e., {\color{blue}$\mu_\xi$} generates {\color{orange}$\underline\xi$}};
		\end{scope}
	\end{tikzpicture}

		\footnotesize		(\emph{moment map:} ${\color{blue}\mu:M\to\g^*}$, 
	 \emph{Hamiltonian $G$-space:} 	$(M,\omega,G,\mu)$)
	\end{center}
	\item
	 $\alpha\mapsto X_\alpha$ is a homomorphism of Leibniz algebras:
	$$\d\{\alpha,\beta\} = \d\L_{X_\alpha}\beta = \L_{X_\alpha}\iota_{X_\beta}\omega = \iota_{[X_\alpha,X_\beta]}\omega ~{\color{black!80} \implies}~
	X_{\{\alpha,\beta\}} = [X_\alpha,X_\beta]$$
}
%------------------------------------------------------------------------------------------------

\subsection{Regular reduction}
%------------------------------------------------------------------------------------------------
\begin{frame}{Regular reduction in symplectic geometry}
	\textbf{\color{UniGreen}Symplectic reduction:}~~
	\\ 
	{\it \small
	Procedure associating to any (suitably regular) pair of symplectic manifold and Hamiltonian action another symplectic manifold of smaller dimension.}
	\vfill
	\pause
	\begin{thmblock}[Marsden-Weinstein reduction \cite{MarsdenWeinstein74}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
		    Given: & $(M,\omega)$ symplectic
		    \\
		    & $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
		    \\[.2em]
		    Assume: & $\phi \in \mathfrak{g}^*$ regular value of $\mu$ 
		    \qquad\quad \footnotesize \textcolor{gray}{($\Rightarrow$ $\mu^{-1}(\phi)\hookrightarrow M$ smooth embedding)}
		    \\
			& $G_\phi\curvearrowright \mu^{-1}(\phi)$ free and proper
			\quad \footnotesize \textcolor{gray}{($\Rightarrow$ $\mu^{-1}(\phi)/G_\phi$ smooth manifold)}
			\\[.4em]
			Then: & $\exists!$ symplectic structure $\omega_\phi$ in $M_\phi:= \mu^{-1}(\phi)/G_\phi$ \\
			& s.t. $\pi^\ast \omega_\phi = j^\ast \omega$ 
			\qquad {\footnotesize with $j:\mu^{-1}(\phi) \hookrightarrow M$ and $\pi:\mu^{-1}(\phi)\twoheadrightarrow M_\mu$}
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}
	%
	\vfill
	\pause
	\begin{columns}
		\begin{column}{0.40\textwidth}
			\includegraphics[width=\textwidth]{Pictures/Reduction}
		\end{column}

		\begin{column}{0.6\textwidth}
				\textbf{\color{UniGreen}In mechanics:}~~
				\\
			{\it \small
				it embodies the process of restricting the dynamics of the system to the level sets of the conserved quantities pertaining to the symmetry group.		
			}
			\\[.2em]
			\color{gray}\small( e.g. restricting to studying a point-like particle in a central potential by studying it in radial coordinates)
		\end{column}	
	\end{columns}	
\end{frame}
\note[itemize]{
	\item {Symplectic Reduction} takes two inputs:
	\\ 1. Hamiltonian $G$-space $(M,\omega,{\color{orange}G},{\color{blue}\mu})$
	\\ 2. parameter ${\color{blue}\phi}\in\g^*$
	\item The \emph{reduced space} is $M_\phi:={\color{blue}\mu^{-1}(\phi)}{\color{orange}/G_\phi}$.
	\item (Marsden--Weinstein '74, Meyer '73) THM:\\
		If $\mu^{-1}(\phi)\subset M$ is smooth and $G_\phi\curvearrowright \mu^{-1}(\phi)$ is free and proper, then there is a \textbf{unique symplectic form} $\omega_\phi\in\Omega^2(M_\phi)$ such that $\pi^*\omega_\phi = i^*\omega$.

	\begin{center}
	\begin{tikzpicture}[scale=2]
		\node[blue] (A) at (0,0) {$\mu^{-1}(\phi)$};
		\node[gray] (A1) at (0,.3) {$i^*\omega$};
		\node[gray] (A2) at (-.6,0) {$\pi^*\omega_\phi$};
		\node[blue] (B) at (1,0) {$M$};
		\node[gray] (B1) at (1,.3) {$\omega$};
		\node[orange] (C) at (0,-.8) {$M_\phi$};
		\node[gray] (C2) at (-.6,-.8) {$\omega_\phi$};

		\path[right hook->,blue] (A) edge node[above] {$i$} (B);
		\path[orange,->] (A) edge node[left] {$\pi$} (C);

		\begin{scope}[xshift=-1.8cm]
			\node[blue] at (0,0) {\small restrict to $\{\mu=\phi\}$};
			\node[orange] at (.15,-.4) {\small quotient by $G_\phi$};
		\end{scope}
	\end{tikzpicture}
	\end{center}
	\item Heuristic Approach to Reduction:
	\\1. describe $G\curvearrowright M$ in terms of $\omega$ 	\hspace{1cm}
	({\color{green}moment map $\mu$})
	\\2. identify a distinguished reduced space \hspace{1cm}
	({\color{green}reduction at $\mu=0$})
	\\3. use the ambiguity in 1.\ to obtain a family of reduced spaces \hspace{1cm}
	({\color{green}reduction at $\mu-\phi=0$, i.e.\ reduction at $\mu=\phi$}
	\\
	{\tiny Note: If $G_\phi\neq G$, then $\mu-\phi:M\to\g^*$ is not a moment map for either $G\curvearrowright M$ or $G_\phi\curvearrowright M$.}

}
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Singular reduction}
\begin{frame}{Symplectic singular reduction schemes}
	\begin{block}{The gist of singular reduction}
 		\begin{itemize}
 			\item[-] when $\mu$ is singular (i.e. $\mu^{-1}(0)$ is not a mfd.), the (geometrically) reduced space may not exist.
 			\item[-] a \emph{singular reduction scheme} is a procedure to construct a "reduced" algebra of observable out of the given data
 			\item[-] such that it corresponds to the algebra of observable of the reduced manifold in the regular case.
 		\end{itemize}
	\end{block}
 	%
 	\pause
 	%
	\begin{thmblock}[Sniatycki-Weinstein reduction \cite{SniatyckiWeinstein83}]
		\vspace{-.4em}\hspace{-1em}
		\begin{tabular}{l p{14cm}}
		    Given: & $(M,\omega)$ symplectic
		    \\
		    & $G\curvearrowright M$ symplectic with equivariant momap. $\mu:M\to \mathfrak{g}^*$
			\\[.4em]
			Then: & 
			$\displaystyle \left[\sfrac{C^\infty(M)}{I_\mu}\right]^G$
			admits a Poisson algebra structure 
			\blfootnote{$I_\mu$ = associative ideal generated by $\widetilde{\mu}(\g)$}			
			\\
			& it agrees with the M--W reduction in the regular case.		
		\end{tabular}
		\vspace{-.4em}
	\end{thmblock}

\end{frame}
\note[itemize]{
 \item Let $(M,\omega,G,\mu)$ be a symplectic Hamiltonian $G$-space.
 \item The \emph{momentum ideal} is the associative ideal $I_\mu\subset C^\infty(M)$ generated by the momenta $\mu_{\xi}$ for any $\xi\in \mathfrak{g}$.
		Namely
		\[
			I_\mu 
			= 
			\Big\langle \mu_\xi \Big\rangle_{\xi \in \mathfrak{g}}^{\text{\tiny asso.}} 
			=
			\left\{
				\left.
					\sum_{i=1}^n f_i ~ \mu_{\xi_i}
				\right|~
					f_i \in C^\infty(M),~ \xi_i\in \mathfrak{g},~  1 \leq i \leq n
			\right\}
		\]

 \item The \emph{\'Sniatycki--Weinstein reduction} is the Poisson algebra
		\[
			\left(C^\infty(M)/I_\mu\right)^G.
		\]

}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Multisymplectic Geometry}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{From symplectic to {multi}symplectic} 
	%
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\pause
		\begin{table}
			\begin{tabular}{c c c}
				symplectic forms \small($n=1$) & $\leftrightsquigarrow$ & volume forms \small($n= \text{dim}(M)-1$)
			\end{tabular}
		\end{table}

	\vfill
	\pause
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
	\end{block}
	%
	\begin{table}
		\ifHandout
		%
		\else
		\only<4>{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map &  
            \\
            \hline
  \multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &  \multicolumn{1}{@{}c@{}}{}              \\
		\end{tabular}
		}
		\fi
		\onslide<5->{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold & multisymplectic manifold \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra & $L_\infty$-algebra \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map & group actions admitting (homotopy) comomentum map
            \\
            \hline
  			\multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &
            \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{field-theoretic systems}}$} 
               \\
		\end{tabular}
		}
	\end{table}

\end{frame}
\note[itemize]{
	\item Historically, the interest in multisymplectic manifolds, has been motivated by the need for understanding the geometrical foundations of first-order classical field theories.
	The key point is that, just as one can associate a symplectic manifold to an ordinary classical mechanical system (e.g. a single
point-like particle constrained to some manifold), it is possible to associate a multisymplectic
manifold to any classical field system (e.g. a continuous medium like a filament or a fluid). See frame Extra-\ref{Frame:Ms-Field-Mechanics}

	\item General ideas basic parallelisms with caveats
	\item caveat: points in multiphase spaces are not states
	\item the table hides the duality between geometric and algebraic approaches to the problem.
	\item	Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[-] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[-] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[-] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Observables in $n$-plectic geometry}
	%
	\begin{defblock}[Hamiltonian $(n-1)$-forms]
		\begin{displaymath}
			\Omega^{n-1}_{ham}(M,\omega) 	:=
			\biggr\{ \sigma \in  \Omega^{n-1}(M) \; \biggr\vert \; 
				\exists \vHam_\sigma \in \mathfrak{X}(M) ~:~ 
				\tikz[baseline,remember picture]{\node[rounded corners,
                        fill=orange!5,draw=orange!30,anchor=base]            
            			(target) {$d \sigma = -\iota_{\vHam_\sigma} \omega$ };
            	}				
				~\biggr\} 
			\end{displaymath}
	\end{defblock}

	\pause
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=orange!5,draw=orange!30,anchor=base]
            	 (base) at ($(current page.north east)-(2,1)$) [rotate=-0,text width=3.5cm,align=center] {\footnotesize{\textcolor{red}{Hamilton-DeDonder-Weyl \\equation}}};
		}	
	\begin{tikzpicture}[overlay,remember picture]
    	\path[->] (base.south) edge[bend right,red](target.north);
    \end{tikzpicture}
	%
	\vfill
	\begin{columns}[T]
		\pause
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\vspace{-1em}
			\begin{thmblock}[Observables $L_\infty$-algebra]
				$\Omega^{n-1}_{ham}(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
				\end{displaymath}			
				can be "completed" to a \\ $L_\infty-algebra$.
			\end{thmblock}
			%	
		\end{column}	
		%
		\pause
		\begin{column}{.50\linewidth}
		%
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\xmark] multiplication of observables;
				\item[\xmark] Jacobi equation;
				%\\ \hspace*{4.25em} full-fledged Jacobi equation;
				\item[\smark] Jacobi equation \emph{up to homotopies}.
			\end{itemize}			
			%
				\[
					\small
					\{\alpha,\{\beta,\gamma\}\} + \text{\it cyc. perm.}
					= {\color{orange}\d \iota_{X_\alpha} \iota_{X_\beta} \iota_{X_\gamma}\omega}
				\]
		\end{column}	
	\end{columns}
	%
	\vfill
	\pause
	Interesting alternatives:
	\begin{itemize}
		\item  descend to $\Omega^{k-1}_{ham}(M) / {\color{orange}\d\hspace{1pt}\Omega^{k-2}(M)}$, hence getting a Lie algebra;
		\item Incorporate the {\color{orange}discrepancy} as part of the data of the space of observables $ 
				L_\infty(M,\omega) = \Omega_{ham}^{k-1}(M)\oplus{\color{orange}\Omega^{\leq k-2}(M)}$
			%a \textbf{homotopy Lie algebra} or \textbf{$L_\infty$-algebra}.
	\end{itemize}

\end{frame}
\note[itemize]{
	\item
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{$L_\infty$-algebra of Observables}
\begin{frame}[fragile,t]{$L_\infty$-algebra of Observables (higher observables) }
	Let be $(M,\omega)$ a $n$-plectic manifold.
	\begin{defblock}[$L_\infty$-algebra of observables ~\emph{(Rogers)} ~\cite{Rogers2010}]

		\hspace{.25em} $L\infty(M,\omega)$ is given by:

		\begin{itemize}
			\item[•] a cochain-complex $(L,\{\cdot\}_1)$ 
		\end{itemize}
		\begin{center}
		\ifHandout
			\includestandalone{Pictures/Figure_Observables}	
		\else
			\includestandalone{Pictures/Frame_Observables}
		\fi				
		\end{center}
		\onslide<2->{
			\begin{itemize}
				\item[•] with $n$ (skew-symmetric) multibrackets $(2 \leq k \leq n+1)$
			\end{itemize}
			\begin{center}
				\includestandalone{Pictures/Equation_Multibracket}	
			\end{center}
		}
		%
	\end{defblock}
  \end{frame}
 \note[itemize]{
	\item if symplectic manifolds are the symmetric take on mechanics, Poisson algebras are the algebraic counterpart.
 	\item A Lie algebra is associated to an ordinary symplectic manifold (its Poisson algebra).
	%(Underlying this is a Lie algebra, whose Lie bracket is the Poisson bracket.)
	Similarly, one associates an Lie-$n$ algebra to any $n$-plectic manifold.
 	% https://ncatlab.org/nlab/show/n-plectic+geometry 	 
 	 %https://ncatlab.org/nlab/show/Poisson+bracket+Lie+n-algebra
	 \item Basically, the higher observables algebra is a chunk of the de Rham complex of $M$ with inverted grading( convention employed here) and an extra structure called "multibrackets".
 	\item ( In the 1-plectic case it reduces to the corresponding Poisson algebra of classical observables)
 	\item Rogers associated to any n-plectic mfd a $L\-\infty$ algebra, Zambon generalized it to the pre-n-plectic case.
 	\item Recognize in the definition of $\{\cdot,\ldots,\cdot\}_k$ the contraction with hamiltonian fields $v_\sigma$ w.r.t. $\sigma$.
  	\item Note $	\iota_{v_{\sigma_1}}\cdots\iota_{v_{\sigma_k}} = (-)^{(k-1)+(k-2)+\dots+1}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}} = (-)^{\frac{k(k-1)}{2}}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}}$ 
 	The definition usually find in literature of Rogers multibrackets involves the coefficient $ (-)^{\frac{k(k-1)}{2}} = -\varsigma(k-1) = (-)^{k+1} \varsigma(k)$.
  \item higher observables is Special instance of a more general object  called $L\-\infty$ Algebra...
 }
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Multisymplectic observables reduction}
\checkpoint	
%-------------------------------------------------------------------------------------------------------------------------------------------------
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Singular Reduction: Roadmap}
	%
	\begin{block}{Data:}
			\begin{itemize}
				\item A \emph{constraint set} $N$ (possibly singular),
				\item An infinitesimal action preserving $N$.
			\end{itemize}
	\end{block}
	%
	\vfill
	\pause
	%
	\begin{block}{Goal:}
		\begin{itemize}
			\item Obtain a "reduced" observables algebra out of the data.
		\end{itemize}
	\end{block}
	%
	\vfill
	\pause
	%
	\begin{block}{Strategy:}
		\begin{enumerate}
			\item Define smooth fields/forms \emph{tangent to $N$},
			\item define smooth fields/forms \emph{vanishing along $N$},
			\item define \emph{reducible fields} requiring the preservation of the vanishing objects,
			\item define \emph{reducible forms} requiring their conservation w.r.t. the infinitesimal action,
			\item define \emph{reducible and vanishing observables},
			\item \textbf{quotient}
		\end{enumerate}
	\end{block}

\end{frame}
\note[itemize]{
	\item
}

%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[shrink]{Smooth objects on a singular set}
	Consider $N$ closed subset of $M$.
	\vfill
	\pause
	\begin{defblock}
	 $I_N$ = ideal of smooth functions vanishing over $N$.
	\end{defblock}
	\vfill
	\pause

	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.65\linewidth}
			%
			\centering \it
				\begin{defblock}[v.f tangent to $N$]
					\begin{displaymath}
						\X_N(M):=
						\left\lbrace
							v \in \X(M)
						~\Big\vert~
							\L_v(I_N) \subseteq I_N
						\right\rbrace
					\end{displaymath}
				\end{defblock}
				\begin{defblock}[v.f vanishing on $N$]
					\begin{displaymath}
						I_\X(N):=
						\left\lbrace
							v \in \X(M)
						~\Big\vert~
							\L_v(C^\infty(M)) \subseteq I_N
						\right\rbrace
					\end{displaymath}				
				\end{defblock}				
		\end{column}	
		%
		%
		\begin{column}{.35\linewidth}
			\centering 
			\includestandalone[width=0.8\textwidth]{Pictures/Figure_VfTangentN}			
		\end{column}	
	\end{columns}			
	%	
	\pause

				%

				%	
		\begin{tcolorbox}[
		enhanced,frame hidden,borderline={0.5pt}{0pt}{blue},
		arc=5pt,colback=white,
		colbacktitle=white,]
			\color{blue}{\textbf{Lem}:} If $N$ is smoothly embedded,  $\X(N)\cong {X_N(M)}/{I_\X}$.
		\end{tcolorbox}

		\pause

		\begin{defblock}[Differential form vanishing on $N$]
			\begin{displaymath}
				I_{\Omega(N)}:=
				\left\lbrace
					\alpha\in\Omega^k(M
				~\left\vert~
					\begin{array}{l r}
						k\geq 0,~		\\		
						\alpha(u_1,\ldots,u_k) \in I_N & \forall u_i \in\X_N(M)
					\end{array}
				\right\rbrace\right.
			\end{displaymath}
		\end{defblock}

\end{frame}
\note[itemize]{
 \item $N$ is not a submanifold in general. An example is $N=\mu^{-1}(0)$ for a certain smooth map $N$.
 \item observe that if $N$ is a smooth embedded submanifold, one has that $\X(N)\cong \frac{X_N(M)}{I_\X}$

}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Multisymplectic singular reduction}
\begin{frame}{Reducible smooth objects \quad \small (w.r.t. $N$ and $\g\action M$)}
	Consider $\g \action M$ by vector fields tangent to $N$ % \hfill( fond. distribution $\underline{\g}\subseteq \X_N(M)$)
	\\
	\vfill
	Denote by :  
	\hspace{1em} $\underline{\g}\subseteq \X_N(M)$ the fundamental distribution,
	\\
	\hspace{6.5em}  $\X_g$ the $C^\infty$-module generated by $\underline{\g}$.
	\\
	\vfill
	\begin{defblock}[Reducible v.fields ]
			\begin{displaymath}
				\X(M)_{[N]} :=
				\left\lbrace
					v \in \X(M)
				~\left\vert~
					\begin{array}{l}
						\L_v (I_N) \subseteq I_N	\\		
						\L_v (\X_g) \subseteq \X_g + I_\X
					\end{array}
				\right\rbrace\right.
			\end{displaymath}
%			\blfootnote{
%			 $\X_g$ = $C^\infty(M)$-module generated by the fundamental distribution.
%			}

	\end{defblock}	
	%
	\pause
	%
	\begin{defblock}[Reducible forms ]
		\begin{displaymath}
			\Omega(M)_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)
			~\left\vert~
				\begin{array}{l r}
					\L_\underline{\xi} \,\alpha \in I_{\Omega(N)}	\\		
					\iota_\underline{\xi} \,\alpha \in I_{\Omega(N)}	& \forall \xi \in \g				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}	
	%
	\pause
	\vfill
	%
	Let $(M,\omega)$ be a $k$-plectic manifold
	\begin{defblock}[Reducible Hamiltonian forms]
		\begin{displaymath}
			(\Omega(M)_{ham}^{n-1})_{[N]} :=
			\left\lbrace
				\alpha \in \Omega(M)_{ham}^{n-1}
			~\left\vert~
				\begin{array}{l r}
					\alpha \text{ is a reducible form} \\
					\vHam_\alpha \text{ is a reducible v.field}
				\end{array}
			\right\rbrace\right.
		\end{displaymath}	
	\end{defblock}

\end{frame}
\note[itemize]{
 \item $\g\action M$ by vector field tangent to $N$ means that $\underline{\xi}\in\X_n(M) \forall \xi \in \g$
 \item Spelling out the definition: reducible vector fields are 
 \\i) v.f. tangent to N 
 \\ii) such that their commutator with any $\underline{\xi}$ lies in $\X_\g$ along $N$.
 \item more algebraically, they stabilize both $I_N$ and $\X_g + I_\X$.
 \item observe that $\L_v I_\X \subseteq I_\X$ since , $\forall u \in I_\X$ $\forall f \in C^\infty(M)$ one has\\
 $(\L_v u) f = \L_{[v,u]} f = \L_v\L_u f - \L_u\L_v f \in I_N$
 \item
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}\frametitle{Hierarchy of auxiliary constructions}
	\begin{minipage}{\textwidth}
	\centering
	\scalebox{.65}{
	\begin{tikzpicture}
		\node (A) at (0,10.0) {\fbox{\parbox{5.6cm}{\centering Ideal of functions vanishing on $N$ \\ $I_N$}}};
		\node (B) at (0,8.3) {\fbox{\parbox{4.4cm}{\centering Vector fields tangent to $N$ \\ $\X_N(M)$}}};
		\node (C) at (3,6.5) {\fbox{\parbox{4.2cm}{\centering Forms vanishing along $N$ \\ $I_\Omega(N)$}}};
		\node (D) at (3,4.2) {\fbox{\parbox{2.7cm}{\centering Reducible forms \\ $\Omega(M)_{[N]}$}}};
		\node (E) at (-3,6.3) {\fbox{\parbox{5.3cm}{\centering Fundamental $C^\infty$-module \emph{plus} \\ vector fields vanishing along $N$ \\ $\X_\g(M)+I_\X(N)$}}};
		\node (E') at (-3,4.2) {\fbox{\parbox{3.8cm}{\centering Reducible vector fields \\ $\X(M)_{[N]}$}}};
		\node (F) at (0,2.4) {\fbox{\parbox{3.7cm}{\centering Reducible observables \\ $L_\infty(M,\omega)_{[N]}$}}};
		\node (G) at (0,0.6) {\fbox{\parbox{5.0cm}{\centering Vanishing ideal of observables \\ $I_{L_\infty}(N)$}}};
		\node (H) at (0,-1.05) {\fbox{\parbox{4.3cm}{\centering Observable reduced space \\ $L_\infty(M,\omega)_N$}}};

		\draw[->] (A)-- node[right=1pt]{{\scriptsize used to define}} (B);
		\draw[->] (B)-- (C);
		\draw[->] (C)-- node[right=.25cm]{\rotatebox{-90}{\small $\subset$}} (D);
		\draw[->] (B)-- (E);
		\draw[->] (D)-- (F);
		\draw[->] (E)-- node[left=.25cm]{\rotatebox{-90}{\small $\subset$}} (E');
		\draw[->] (E')-- (F);
		\draw[->] (F)-- node[right=.25cm]{\rotatebox{90}{\small $\subset$}} (G);
		\draw[->] (G)-- (H);

		\node[right=2.95cm] at (A) {\scriptsize\emph{associative ideal}};
		\node[right=2.30cm] at (B) {\scriptsize\emph{Lie algebra}};
		\node[right=2.25cm] at (C) {\scriptsize\emph{differential ideal}};
		\node[right=1.5cm] at (D) {\scriptsize\emph{differential graded subalgebra}};
		\node[left=2cm] at (E') {\scriptsize\emph{Lie subalgebra}};
		\node[left=2.8cm] at (E) {\scriptsize\emph{Lie ideal}};
		\node[right=2cm] at (F) {\scriptsize\emph{$L_\infty$-subalgebra}};
		\node[right=2.65cm] at (G) {\scriptsize\emph{$L_\infty$-ideal}};
		\node[right=2.3cm] at (H) {\scriptsize\emph{$L_\infty$-algebra}};
	\end{tikzpicture}
	}
	\end{minipage}
\end{frame}
\note[itemize]{
 \item
}

%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Multisymplectic observable reduction}
	\begin{defpropblock}[Reducible $L_\infty$-observables]
		Is the {\color{blue!70!black}$L_\infty$-subalgebra} of $L_\infty(M,\omega)$ given by
		\begin{displaymath}
			L_\infty(M,\omega)_{[N]}^k :=
			\begin{cases}
				\Omega^{n-1-k}(M)_{[N]} 
				\qquad\text{\color{gray}\small (reducible forms) }
				& \text{if } n-1\leq k < 0 \\
				(\Omega(M)_{ham}^{n-1})_{[N]} 
				\quad
				\text{\color{gray}\small (reducible hamiltonians) }
				& \text{if } k = 0 \\
				0 & \text{if } k > 0
			\end{cases}
		\end{displaymath}
	\end{defpropblock}
	%
	\pause
	%
	\begin{defpropblock}[Vanishing $L_\infty$-observables]
		Is the {\color{blue!70!black}$L_\infty$-ideal} of $L_\infty(M,\omega)_{[N]}$ given by
		\begin{displaymath}
			I_{L_\infty(M,\omega)} :=
			\left\lbrace
				\alpha \in L_\infty(M,\omega)_{[N]}
			~\left\vert~
				\begin{array}{l l}
					\alpha(v_1,\dots,v_k) \in I_N  \quad \forall v_i \in \X_N &
					\text{if}~ \alpha \in \Omega^k \\
					\vHam_\alpha \in \X_\g + I_\X &
					\text{if}~ \alpha \in \Omega^{n-1}
				\end{array}
			\right\rbrace\right.
		\end{displaymath}
	\end{defpropblock}
	%
	\pause
	%
	\begin{thmblock}[Blacker - M. - Ryvkin]%[Reduced $L_\infty$-algebra of observables]
		The reduction of $L_\infty(M,\omega)$  with respect to $\g\curvearrowright (N\subset M)$ is the %$L_\infty$-algebra
		quotient
		\begin{displaymath}
					L_\infty(M,\omega)_N = \frac{L_\infty(M,\omega)_{[N]}}{I_{\L_\infty}(N)}~.		
		\end{displaymath}
	\end{thmblock}

\end{frame}
\note[itemize]{
 \item The graded vector space underlying the reduced $L_\infty$-algebra $\Ham_\infty(M,\omega)_N$ is explicitly given by
\begin{displaymath}
	\Ham_\infty(M,\omega)_N =
	\frac{
		\left\lbrace
			(\alpha,v) \in \Omega(M){[n{-}1]}\oplus\X(M)
		~\left\vert~
			\begin{array}{l l}
				\iota_v \omega = -\d \alpha^{(0)} &
				\\
				\iota_\xi \alpha \in I_{\Omega}(N) &
				\\
				\L_\xi\alpha \in I_{\Omega}(N)
				&
				\\
				\L_\xi v \in \fgmodule +\vanvf
				&~\forall \xi \in \g 
				\\
				v \in \X_N(M)
				&
			\end{array}
		\right.
		\right\rbrace
	}{
		\left\lbrace
			(\alpha,v) \in \Omega(M){[n{-}1]}\oplus\X(M)
		~\left\vert~
			\begin{array}{l}
				\alpha \in I_{\Omega}(N)
				\\
				v \in \fgmodule +\vanvf
			\end{array}
		\right.
		\right\rbrace
	}~.
\end{displaymath}
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}{Singular reduction (Conclusions)}
		  \centering 
	\emph{Reduction of states vs.\ reduction of observables}
	\vfill
	\begin{minipage}{.7\textwidth}
	\hspace{-.4cm}
	\begin{tikzpicture}
		\draw (-4,-2.5) rectangle (4,2.5);
		\draw (-4,0) -- (4,0);
		\draw (0,-2.5) -- (0,2.5);

		\draw node[align=center] at (-2,1.25) {symplectic\\ reduction};
		\draw node[align=center] at (2,1.25) {multisymplectic\\ reduction};
		\draw node[align=center] at (-2,-1.25) {\'{S}niatycki--Weinstein,\\Dirac,\\Arms--Gotay--Jennings\\ reduction};
		\draw node[align=center] at (2,-1.25) {\color{UniGreen}$L_\infty$ reduction};

		\draw node[align=center] at (2,2.85) {\color{UniGreen}\emph{higher}};
		\draw node[align=center] at (-2,2.85) {\emph{lower}};

		\draw node[align=right] at (-4.7,1.25) {\emph{states}};
		\draw node[align=right] at (-5.13,-1.25) {\color{UniGreen}\emph{observables}};
	\end{tikzpicture}
	\end{minipage}

	\pause
			\vfill
		  \centering 
		  {\Huge\color{red} 
		  \emph{Thank you for your attention!}}
\end{frame}
\note[itemize]{
 \item Consider $N=\mu^{-1}(0)$ to be regular (smooth embedding)
		\vspace{.5em}
		\begin{table}[]
			\begin{tabular}{ccc}
			Multisymplectic               &                      & Multisymplectic                  \\
			regular                       & $\equiv$             & singular                \\
			reduction & & reduction
			\end{tabular}
		\end{table}

		\item
			Consider $\omega$ to be $1$-plectic
			\vspace{.5em}
			\begin{table}[]
				\begin{tabular}{ccc}
				Multisymplectic               &                      &  Sniaticky--Weinstein                  \\
				singular                       & $\cancel\equiv$             & singular                \\
				reduction & & reduction
				\end{tabular}
			\end{table}

			(but $\exists$ a canonical Poisson algebra morphism)

}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
% APPENDIX
%------------------------------------------------------------------------------------------------
\appendix
%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Complementary Material}
%##################################################################################
\begin{frame}
	\begin{center}
	\Huge\emph{Supplementary Material}
	\end{center}
\end{frame}
\note[itemize]{
	\item
}
\addtocounter{framenumber}{-1}
%##################################################################################

%------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{MS geometry and classical field mechanics}\label{Frame:Ms-Field-Mechanics}
		Consider a smooth manifold $Y$,
		\begin{columns}
			\hfill
			\begin{column}{.5\linewidth}
				\emph{Multicotangent bundle} $\bigwedge = \bigwedge^n T^\ast Y$\\
				is naturally $n$-plectic
			\end{column}
			\begin{column}{.4\linewidth}
				\[
				\begin{tikzcd}
					\Lambda \ar[d,"\pi"'] & T \Lambda \ar[d,"T \pi"] \ar[l] \\
					Y								& T Y \ar[l]
				\end{tikzcd}	
				\]
			\end{column}
		\end{columns}
	\pause
	\begin{defblock}[Tautological $n$-form]
		$\theta \in \Omega^n(\Lambda)$ such that:
		\begin{displaymath}
		\begin{split}
			\left[ \iota_{u_1 \wedge \ldots \wedge u_n} \theta \right]_\eta 
			&= \iota_{(T \pi)_\ast u_1 \wedge \ldots \wedge (T \pi)_\ast u_n} \eta \\
			&= \iota_{u_1 \wedge \ldots \wedge u_n} \pi^\ast \eta 
			\qquad \qquad \forall \eta \in \Lambda \, , \: \forall u_i \in T_\eta \Lambda 		
		\end{split}
		\end{displaymath}
	\end{defblock}
	\vfill
	\begin{columns}
		\begin{column}{.6\linewidth}
			\begin{defblock}[Tautological (multisymplectic) (n+1)-form]
				$$\omega := d \theta$$
			\end{defblock}
		\end{column}
		\begin{column}{.4\linewidth}
		 	\begin{claimblock}$\omega$ is not degenerate.\end{claimblock}	
		\end{column}
	\end{columns}	
	\pause
	\begin{keywordblock}
		\begin{tabular}{|c|c|c|}
			\hline 
			point-particles mechanics & $\rightsquigarrow$ & classical fields mechanics \\
			%(finite discrete DOF) & & (finite dimensional continuous DOF) \\
			\hline 
			symplectic & $\rightsquigarrow$ & multisymplectic \\ 
			\hline 
			Observables (Poisson) algebra & $\rightsquigarrow$ & Observables $L-\infty$ algebra
			 \\ 
			\hline 
			Co-moment map & $\rightsquigarrow$ & Homotopy co-momentum map \\ 
			\hline 
		\end{tabular} 
	\end{keywordblock}

\end{frame}
\note[itemize]{
	\item This example is significant from the perspective of geometric classical field theory:
		\begin{displaymath}
			\frac{\text{classical mechanics}}{\text{symplectic geo.}} =
			\frac{\text{classical field mechanics}}{\text{multisymplectic geo.}}
		\end{displaymath}
	\item Multicotangent bundle is the \emph{Higher analogue} of the cotangent bundle.
	(but it is not yet the analogue of a \emph{phase space}.)
\item The multiphase space is the sub-bundle of $n$-forms vanishing when contracted with 2 vertical fields.
  	\item The reason why this sub-bundle has a particular role is that it can be proved to be isomorphic to a suitable dual of the first Jet bundle.
  	\item For further details see Gotay et al. \href{https://arxiv.org/abs/physics/9801019}{arXiv:physics/9801019}. For a pictorial representation of all the structures involved in the geometric mechanics of I order classical field theories see appendix, pag: \ref{frame:Gimmsy}.
}
%------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t]{Symmetries in \textbf{multisymplectic geometry}}
	Consider a Lie algebra action $v:\mathfrak{g} \to \mathfrak{X}(M)$  preserving the $n$-plectic form $\omega$,
	\vfill

	\vspace{-1em}
	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			\onslide<2->{
				$-$ symplectic case $-$
				\begin{defblock}[Comoment map pertaining to $v$]
					Lie algebra morphism
					$$ f: \mathfrak{g} \to C^\infty(M) $$
					such that
					$$ d~f (x) = -\iota_{v_x} \omega \qquad \forall x \in \mathfrak{g}~.$$
				\end{defblock}
			}
		\end{column}	
		%
		\onslide<2->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\centering \it
			\onslide<3->{			
				$-$ $n$-plectic case $-$
				\begin{defblock}[Homotopy comoment map \tiny (HCMM)]
					$L_\infty$-morphism 
					$$ (f_k) : \mathfrak{g} \to L_\infty (M,\omega)$$
					such that
					$$ d~f_1(x) = -\iota_{v_x} \omega \qquad \forall x \in \mathfrak{g}~.$$
				\end{defblock}	
			}
		\end{column}	
	\end{columns}	
	%
	\pause
	\vfill
	\centering 
	\onslide<4->{\textbf{-- Conserved quantities --}}
	%
	\vspace{-.5em}
	\begin{columns}[T]
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\centering \it
			\onslide<4->{
			\begin{propblock}[Noether Theorem]
				\small Fixed $H\in C^\infty_{\text{Ham}}(M)$ ($\mathfrak{g}$-invariant) ,
				$$\mathcal{L}_{v_H} f(x) = 0 \qquad \forall x \in \mathfrak{g}$$
			\end{propblock}
			}
		\end{column}	
		%
		\onslide<5->{\vrule{}}
		%
		\begin{column}{.50\linewidth}
			\centering \it
			\onslide<5->{			
			\begin{propblock}[RWZ16 Theorem]
				\small Fixed $H\in \Omega^{n-1}_{\text{Ham}}(M)$ ($\mathfrak{g}$-invariant),
				$$\mathcal{L}_{v_H} f_k(p) \in B^k(M) \qquad \forall p \in Z_k(\mathfrak{g})$$			
			\end{propblock}
			}
		\end{column}	
	\end{columns}		
\end{frame}
\note{
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\begin{frame}{Reminder: $L_\infty$ Algebras}

		\emph{
			$L_\infty$-algebra is the notion that one obtains from a Lie algebra when one requires the Jacobi identity to be satisfied only up to a higher coherent chain homotopy.
		}
		\\
		\vspace{.5em}
		\begin{defblock}[$L_\infty$-algebra ~\emph{(Lada, Markl)} ~\cite{Lada1995}]
			\includestandalone{Pictures/Figure_Linfinitydef}
		\end{defblock}	
		%
		%
	\pause
	\vfill
	\begin{thmblock}[Rogers \cite{Rogers2010}]
		The \emph{higher observable algebra} $L_{\infty}(M,\omega)$ 	forms an honest $L_\infty$ algebra.
		\footnotetext{ Take $\mu_1 = \text{d}$, $\mu_k=\lbrace\dots\rbrace_k$, $L$ is a shifted truncation of the de Rham complex.}
	\end{thmblock}
\end{frame}
\note[itemize]{
	\item $L_\infty$-algebra is the notion obtained from a Lie algebra requiring that the Jacobi identity is satisfied only up to a higher coherent chain homotopy.
	\item The Lie-n algebra mentioned before is a $L_\infty$ algebra with underlying graded vector space concentrated in degrees $0,1...n$.

	\item Definition. We say that a permutation $\sigma \in S_n$ is a $(j,n-j)$-unshuffle, $0\leq j \le1 n$  if $\sigma(1)< \dots < \sigma(j)$ and $\sigma(j+1)<\dots<\sigma(n)$.
	\\
	You can also say that $\sigma$ is a $(j,n-j)$-unshuffle if $\sigma(i)< \sigma(i+1)$ when $i\neq j$.

	\item 	Alternatively, the Jacobiators can be also denoted as $$\displaystyle J_m=\sum_{i+j=m+1} 	\mu_i \triangleleft \mu_j = 0$$
	employing the so-called \emph{ Richardson-Nijenhuis product}
		 $$\mu_i\circ \mu_j := (-)^{i(j+1)}\frac{1}{j!(i-1)!}\mu_i \triangleleft \mu_j\otimes \mathbb{1}_{i-1} \circ \mathcal{A}~,$$
		 where $\mathcal{A}$ denotes the (graded) total skew-symmetrizator.

	\item see frame extra-\ref{Frame:unwapping-Jacobi} for a slightly demystification of the higher Jacobi equations.

	\item more precisely this statement is a proposition/definition

}
%------------------------------------------------------------------------------------------------

	\input{msred-aknowledgements}
%------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\end{document}